Theoretical prediction of disrupted Min oscillation in flattened \emph{Escherichia coli}

### 2015-06-20 paper/response-to-referees.txt

  <Jeff does this.>

Minor points:

4. The referee points out that MinD clusters had been considered by
Kruse prior to Huang.  We thank the referee for pointing out this
oversight, and have updated our background section to include this
reference and more accurately reflect the historical development.

5. Based on the referee's suggestion, we have clarified that reference
11 does not extend Huang's model, except in that it simulates the same
reaction-diffusion equations stochastically.

@@

6. The referee points out that two previous works have also
demonstrated that stochastic simulations are needed in order to
explain WHAT?.  WE HAVE EDITED OUR DISCUSSION TO ILLUSTRATE THIS.

  <David does this.>

Referee #3.

7. The referee #3 also expressed doubt about our conclusion with
regard to the effect of irregular shape which we base on our stadium
models.  Unlike the previous referee, the criticism is based not on
the spatial output (Figure 4), but rather on the reported coherence
times.  We had attempted to make a distinction between spatially
irregular oscillation and temporally irregular oscillations.  This
distinction was evidently not clear, so we have ADDED AN EXPLICIT
DISCUSSION OF THE FACT THAT WE OBSERVE A GREATER DISRUPTION IN
TEMPORAL REGULARITY IN THE IRREGULARLY SHAPED CELLS.  JEFFSAYS{Changed
the last sentence before the conclusion.  I don't think it needs much
changing except maybe moderate our claims a bit (which I tried to
do).}  FIXME WHAT DO WE DO ABOUT SPATIAL IRREGULARITY?

@@

  <Jeff does this.>

8. The referee also questioned how the cell was divided in order to
perform the coherence time calculations, pointing out that a smaller
region could lead to a shorter coherence time.  EXPLAIN HOW WE DIVIDE
THE CELLS UP IN THIRDS BY VOLUME? IT'S AWKWARD THAT WE DON'T DIVIDE
ALONG LONG AXIS.  JEFFSAYS{Bad news here.  It looks like I did this
dividing up the cells by eye.  Originally the slices were used to
created the box plots, which was only a way to visualize the data in
smaller time domains, so there's wasn't really any disadvantage to
doing it roughly like this (we even had the visually appealing little
colored diagrams on the sides of the box plots).  Not sure what to say
abou this in text.}

### 2015-06-20 paper/plos.tex
more changes on Saturday

\begin{figure*}
  \centering
  \hspace{-3cm}
  \vspace{1cm}
  \includegraphics[width=16cm]{../paper/plot-ave}
  \caption{We display here arrows depicting successive maxima in space
    and and time overlayed on a color plot of the total MinD density
    averaged over the same time period.  The simulation time covered
    for the flattened cells is 350 seconds, which is the same
    period of time depicted in the experimental data plots of M\"annik
    \emph{et al.}~\cite{mannik2012robustness}.  For the wild-type pill
    shape, we only cover 250 seconds, in order to provide a
    useful comparison due to its shorter oscillation period.  The top
    row shows plots published by M\"annik \emph{et al.} of the MinD maxima behavior
    and the bottom two rows show our simulations using the stochastic
    and deterministic models, respectively.  We simulated
    approximations to the two shapes observed by M\"annik, which we call
    \emph{shape A} and \emph{shape B}.  In addition, we studied two
    flattened stadium shapes which we call \emph{stadium A} and
    \emph{stadium B} corresponding in aspect ratio and thickness to
    the two experimental shapes.  The spatial length scale of all the
    figures shown was identical. Each of the flattened cell shapes
    uses the same color scale for the number of proteins per unit
    area.  Finally, we display the natural pill shape, which was also
    featured in Figs.~\ref{image-p} and~\ref{corr-pill}, with a
    different color scale to reflect the thicker cell containing more
    proteins per cross-sectional area.  }
  \label{randst-plot-ave}
\end{figure*}

@@

The predictions of the stochastic model are roughly similar to the
experimental results in terms of irregularity of the locations of
maxima, although there are locations on the edges of the cells where
experiment shows maxima occurring that we never see in our
simulations, suggesting that the model does not precisely reflect the
experiment.  We also note that the stadium shapes in
Fig.~\ref{randst-plot-ave} appear to have maxima location irregularity
\fixme{within a small number of periods} that is
qualitatively similar to our irregular shapes and to experiment.

@@

We see that both in the shape \emph{B} and stadium \emph{B}, the
deterministic model is perfectly periodic.  Both the deterministic and
stochastic model correlations have the same number of proteins in
their cells, and have been normalized with the same normalization
factor, so that they can be directly compared against each other.  The
stochastic simulation of shape \emph{B} has a coherence time of 5.3
periods, while the corresponding simulation of stadium \emph{B} has a
coherence time of 8.6 periods.
%
The correlation function for stadium \emph{A} is similar to that of
stadium~\emph{B}, with similar coherence time (13 periods, not shown),
while shape~\emph{A} also exhibits a similar coherence time (5.5
periods).  This confirms that the stadium shapes do indeed oscillate
somewhat more regularly than the irregular shapes, but both are
comparable to the wild-type pill shape.
%
From the stochastic correlation data, we conclude that irregularity of
shape has only a moderate affect on the temporal regularity of the
oscillatory behavior, with even the very distorted shape~\emph{B}
exhibiting only a factor of two greater decoherence.

@@

We find that flattened but regular and symmetric cells exhibit MinD
oscillations that are qualitatively similar to the oscillations
observed experimentally (and in our simulation) in irregular and
asymmetric cells.  This demonstrates that asymmetry is not required in
order to induce spatially irregular MinD oscillations.  In addition,
the temporal regularity of oscillations is only moderately affected by
irregularity in the shape of flattened cells.  Thus we conclude that
it is the flattening of the cells, and the accompanying expansion in the
transverse direction, rather than their irregular shape that primarily
causes the disruption of regular bipolar oscillation which is observed
by M\"annik \emph{et al.}~\cite{mannik2012robustness}.

>>>>

\begin{figure*}
  \centering
  \hspace{-3cm}
  \vspace{1cm}
  \includegraphics[width=16cm]{../paper/plot-ave}
  \caption{Arrows depicting successive maxima in space
    and time overlayed on a color plot of the total MinD density
    averaged over the same time period.  The simulation time covered
    for the flattened cells is 350 seconds, which is the same
    period of time depicted in the experimental data plots of M\"annik
    \emph{et al.}~\cite{mannik2012robustness}.  For the wild-type pill
    shape, we only cover 250 seconds, in order to provide a
    useful comparison due to its shorter oscillation period.  The top
    row shows plots published by M\"annik \emph{et al.} of the MinD maxima behavior
    and the bottom two rows show our simulations using the stochastic
    and deterministic models, respectively.  We simulated
    approximations to the two shapes observed by M\"annik, which we call
    \emph{shape A} and \emph{shape B}.  In addition, we studied two
    flattened stadium shapes which we call \emph{stadium A} and
    \emph{stadium B} corresponding in aspect ratio and thickness to
    the two experimental shapes.  The spatial length scale of all the
    figures shown was identical. Each of the flattened cell shapes
    uses the same color scale for the number of proteins per unit
    area.  Finally, we display the natural pill shape, which was also
    featured in Figs.~\ref{image-p} and~\ref{corr-pill}, with a
    different color scale to reflect the thicker cell containing more
    proteins per cross-sectional area.  }
  \label{randst-plot-ave}
\end{figure*}

@@

The predictions of the stochastic model are roughly similar to the
experimental results in terms of irregularity of the locations of
maxima, although there are locations on the edges of the cells where
experiment shows maxima occurring that we never see in our
simulations, suggesting that the model does not precisely reflect the
experiment.  We also note that the stadium shapes in
Fig.~\ref{randst-plot-ave} appear to have maxima location irregularity
\fixme{within the time domain of a small number of periods, as
  compared to the more long term coherence discussed below,} that is
qualitatively similar to our irregular shapes and to experiment.

@@

We see that both in the shape \emph{B} and stadium \emph{B}, the
deterministic model is perfectly periodic.  Both the deterministic and
stochastic model correlations have the same number of proteins in
their cells, and have been normalized with the same normalization
factor, so that they can be directly compared against each other.  The
stochastic simulation of shape \emph{B} has a coherence time of 5.3
periods, while the corresponding simulation of stadium \emph{B} has a
coherence time of 8.6 periods.
%
The correlation function for stadium \emph{A} is similar to that of
stadium~\emph{B}, with similar coherence time (13 periods, not shown),
while shape~\emph{A} also exhibits a similar coherence time (5.5
periods).  This confirms that the stadium shapes do indeed oscillate
somewhat more regularly than the irregular shapes, but both are
comparable to the wild-type pill shape.
%
From the stochastic correlation data, we conclude that irregularity of
shape has \fixme{a distinct yet moderate} affect on the temporal regularity of
the oscillatory behavior, with even the very distorted shape~\emph{B}
exhibiting only a factor of two greater decoherence.

@@

We find that flattened but regular and symmetric cells exhibit MinD
oscillations that are qualitatively similar to the oscillations
observed experimentally (and in our simulation) in irregular and
asymmetric cells.  This demonstrates that asymmetry is not required in
order to induce spatially irregular MinD oscillations.  In addition,
the temporal regularity of oscillations is only moderately affected by
irregularity in the shape of flattened cells.  Thus we conclude that
it is \fixme{primarily} the flattening of the cells, and the
accompanying expansion in the transverse direction, \fixme{deleted
  some text} that causes the disruption of regular bipolar oscillation
which is observed by M\"annik \emph{et
  al.}~\cite{mannik2012robustness}.
